---
title: "cheat-sheet"
output: 
  html_document: 
    keep_md: yes
date: "2023-01-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lab 1

*gets current working directory*
```{r}
getwd()
```
*arithmetic order of operations applies*
*help on method*
```{r}
?mean
```

*vectors*
```{r}
example_1 <- c(1,2,3,4,5,6) # c is concatenate
```
- operations on vectors
```{r}
mean(example_1)
```
```{r}
median(example_1)
```
```{r}
sd(example_1)
```


## Lab 2: objects and data classes
*operations to work with objects*
```{r}
#making the objects
red <- 34
yellow <- 45
purple <- 29
```
```{r}
sum(red,yellow)
```

*classes*
```{r}
my_numeric <- 42 
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```

```{r}
#important functions
class(my_character) #class of x
is.integer(my_character) # checks is x is of specified type, Boolean return
new_integer <- as.character(my_integer) # changes class of x to specified class
class(new_integer)
```
*NA values*
```{r}
bleh <- NA
is.na(bleh) # checks if bleh is NA type
anyNA(bleh) #same as above but more useful for vectors, data matrices and data frame
```
```{r}
blehbleh <- c(1,5,2,NA,4,6)
mean(blehbleh, na.rm = T ) #na. tells function whether to include the NA value or not
mean(blehbleh) # will give a null value without it
```
- Data Structures
In addition to classes of data, R also organizes data in different ways. These are called data structures and include vectors, lists, matrices, data frames, and factors. For now, we will focus on `vectors`.  

- Vectors
Vectors are a common way of organizing data in R.  We create vectors using the `c` command. The `c` stands for concatenate. 

```{r}
#creating a sequence
v <- c(1:50)
v
#evaluating
v <= 25
```
```{r}
# to get only elements
v[v<= 25]
```
*accesing elements*
```{r}
days_of_the_week <- c("Monday", "Tuesday", "Wednesday", "Thrusday", "Friday", "Saturday", "Sunday")
days_of_the_week[3]
```

- Data Matrices
Data matrices are a series of stacked vectors, similar to a data table. In the example below, we build a new data matrix using the matrix command.



