---
title: "Lab 4 Homework"
author: "Khushleen Kaur"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  
Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

**1. Load the data into a new object called `homerange`.**

```{r}
homerange <- readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```
**2. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.**  

```{r}
# shows dimension of data
dim(homerange)
```
```{r}
# displays names of all columns in the database
names(homerange)
```
```{r}
# provides a summary of all columns including character type, length and statistical summary
summary(homerange)
```

**3. Change the class of the variables `taxon` and `order` to factors and display their levels.**  

```{r}
# Changing taxon class to factor and display levels
homerange$taxon <- as.factor(homerange$taxon)
levels(homerange$taxon)

```

```{r}
# Changing order class to factor and display levels
homerange$order <- as.factor(homerange$order)
levels(homerange$order)
```

**4. What taxa are represented in the `homerange` data frame? Make a new data frame `taxa` that is restricted to taxon, common name, class, order, family, genus, species.**  

- Taxa represented in homerange are birds, lake fishes, lizards, mammals, marine fished, river fishes, snakes, tortoises and turtles.
```{r}
# Displays all Taxa represented in the homerange data frame.
levels(homerange$taxon)
```
```{r}
# new restricted data frame
taxa <- select(homerange, "taxon":"species")
taxa
```

**5. The variable `taxon` identifies the large, common name groups of the species represented in `homerange`. Make a table the shows the counts for each of these `taxon`.**  

```{r}
table(homerange$taxon)
```

**6. The species in `homerange` are also classified into trophic guilds. How many species are represented in each trophic guild.** 

- There are 342 carnivorous species and 227 herbivorous species.
```{r}
table(homerange$trophic.guild)
```

**7. Make two new data frames, one which is restricted to carnivores and another that is restricted to herbivores.** 

```{r}
carnivores_only <- filter(homerange, trophic.guild == "carnivore")
carnivores_only
```
```{r}
herbivores_only <- filter(homerange, trophic.guild == "herbivore")
herbivores_only
```

**8. Do herbivores or carnivores have, on average, a larger `mean.hra.m2`? Remove any NAs from the data.**  

- Based on the calculations below we can see that herbivores have the larger mean.hra.m2 of the two
```{r}
# carnivore mean
mean(carnivores_only$mean.hra.m2, na.rm = T)
```
```{r}
# herbivore mean
mean(herbivores_only$mean.hra.m2, na.rm = T)
```
**9. Make a new dataframe `deer` that is limited to the mean mass, log10 mass, family, genus, and species of deer in the database. The family for deer is cervidae. Arrange the data in descending order by log10 mass. Which is the largest deer? What is its common name?**  

- The largest deer based on the calculations below is the moose (common name) and belongs to the alces species
```{r}
#filters out only deer
complete_deer <- filter(homerange, family == "cervidae")
# shows only select few columns for the data base
deer <- select(complete_deer, "mean.mass.g", "log10.mass", "family", "genus", "species")
```
```{r}
# arranges the data in decreasing order by log 10 mass
arrange(deer, desc(log10.mass))
```
```{r}
max_mass = max(deer$log10.mass)
# displays information about the largest deer
complete_deer[complete_deer$log10.mass == max_mass, 2]
```


**10. As measured by the data, which snake species has the smallest homerange? Show all of your work, please. Look this species up online and tell me about it!** **Snake is found in taxon column**    

- The snake species with the smallest homerange (as shown below) is Bitis schneideri or as it is commonly called, namaqua dwarf adder. According to wikipedia, this snake has an avverage total length of 18 - 25 cm and is the smallest species in the Bitis genus. It is found between South Africa and the Namibian desert.
```{r}
# filters put only snakes into a new database
snakes <- filter(homerange, taxon == "snakes")
minimum_hra <- min(snakes$mean.hra.m2)
```
```{r}
# outputs information about snake with smallest homerange.
snakes[snakes$mean.hra.m2 == minimum_hra, 2]

```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   